---
title: "P8105 Homework 2"
author: Mengfan Luo (ml4701)
output: github_document
---

```{r, echo=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
```

### Problem 1

Let's load the Mr. Trash Wheel sheet as `mtw_df` and clean it:

```{r}
mtw_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "Mr. Trash Wheel",range = "A2:N534") %>% 
  drop_na() %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls)) 

```

Let's take a look at the first and last 3 lines of `mtw_df`.

```{r}
knitr::kable(head(mtw_df,3))
knitr::kable(tail(mtw_df,3))
```


Let's load and clean precipitation data for 2018 and 2019, and then combine them into one dataframe `prec_2018_2019_df`.

```{r}
prec_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2018 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

prec_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2019 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)

prec_2018_2019_df = 
  bind_rows(prec_2018_df,prec_2019_df) %>% 
  mutate(month = month.name[month])

```


Let's take a look at the first and last 3 lines of `prec_2018_2019_df`.

```{r}
knitr::kable(head(prec_2018_2019_df,3))
knitr::kable(tail(prec_2018_2019_df,3))
```

 
For `mtw_df` which is tidied from Mr. Trash Wheel sheet, there are ``r nrow(mtw_df)`` observations and ``r ncol(mtw_df)`` variables (including serial number of the dumpster `dumpster`). Other variables are ``r names(mtw_df)[-1]``.

For `prec_2018_2019_df` which is combined from cleaned precipitation data for 2018 and 2019, there are there are ``r nrow(prec_2018_2019_df)`` observations and ``r ncol(prec_2018_2019_df)`` variables, inclusing ``r names(prec_2018_2019_df)``. We can calculate from this dataframe that the total precipitation in 2018 was ``r sum(pull(prec_2018_2019_df,total)[1:12])``, and the median number of sports balls in a dumpster in 2019 was ``r median(pull(prec_2018_2019_df,total)[13:24])``.
 

 
 