---
title: "P8105 Homework 2"
author: Mengfan Luo (ml4701)
output: github_document
---

```{r, echo=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

### Problem 1

Let's load the Mr. Trash Wheel sheet as `mtw_df` and clean it:

```{r}
mtw_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "Mr. Trash Wheel",range = "A2:N534") %>% 
  drop_na() %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls)) 

```

Let's take a look at the first and last 3 lines of `mtw_df`.

```{r}
knitr::kable(head(mtw_df,3))
knitr::kable(tail(mtw_df,3))
```


Let's load and clean precipitation data for 2018 and 2019, and then combine them into one dataframe `prec_2018_2019_df`.

```{r}
prec_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2018 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

prec_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2019 Precipitation",range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)

prec_2018_2019_df = 
  bind_rows(prec_2018_df,prec_2019_df) %>% 
  mutate(month = month.name[month])

```


Let's take a look at the first and last 3 lines of `prec_2018_2019_df`.

```{r}
knitr::kable(head(prec_2018_2019_df,3))
knitr::kable(tail(prec_2018_2019_df,3))
```

 
For `mtw_df` which is tidied from Mr. Trash Wheel sheet, there are ``r nrow(mtw_df)`` observations and ``r ncol(mtw_df)`` variables (including serial number of the dumpster `dumpster`). Other variables are ``r names(mtw_df)[-1]``.

For `prec_2018_2019_df` which is combined from cleaned precipitation data for 2018 and 2019, there are there are ``r nrow(prec_2018_2019_df)`` observations and ``r ncol(prec_2018_2019_df)`` variables, inclusing ``r names(prec_2018_2019_df)``. We can calculate from this dataframe that the total precipitation in 2018 was ``r sum(pull(prec_2018_2019_df,total)[1:12])``, and the median number of sports balls in a dumpster in 2019 was ``r median(pull(prec_2018_2019_df,total)[13:24])``.
 
 
### Problem 2

Our goal is to merge these into a single data frame using year and month as keys across datasets.

1. clean and tidy the data in pols-month.csv. First 6 lines of the dataframe will be shown. 

```{r,message=FALSE}
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon,into = c("year","month","day"),sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  pivot_longer(cols = starts_with("prez"),names_to = "president",
               names_prefix = "prez_",values_to = "prez_value") %>% 
  filter(prez_value==1) %>% 
  select(-day,-prez_value) 

knitr::kable(head(pols_df))
```


2. clean, tidy and organize the data in snp.csv. First 6 lines of the dataframe will be shown. 

```{r,message=FALSE}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  mutate(date = mdy(date)) %>% 
  separate(date,into = c("year","month","day"),sep = "-") %>% 
  select(-day) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year > 2049,year -100,year)) %>% 
  arrange(year,month) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month])
knitr::kable(head(snp_df))
```



3. tidy the unemployment data so that it can be merged with the previous datasets. First 6 lines of the dataframe will be shown. 


```{r,message=FALSE}
unemploy_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") 
colnames(unemploy_df)=c("year","1","2","3","4","5","6","7","8","9","10","11","12")

unemploy_df =
  pivot_longer(unemploy_df,`1`:`12`,
               names_to = "month",values_to = "unemploy_percent") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month])
knitr::kable(head(unemploy_df))
```



4. Join the datasets by merging snp into pols, and merging unemployment into the result. First and last 6 rows of the resulted dataframe `merge3_df` are shown below.

```{r}
merge2_df = left_join(pols_df,snp_df,by = c("year","month"))
merge3_df = left_join(merge2_df,unemploy_df,by = c("year","month"))
knitr::kable(head(merge3_df))
knitr::kable(tail(merge3_df))
```


In this problem, 3 dataframe was tidied and merged together.

Dataframe 1 `pols_df` has ``r nrow(pols_df)`` rows and ``r ncol(pols_df)`` columns. The variables in the dataframe are ``r names(pols_df)``.

Dataframe 2 `snp_df` has ``r nrow(snp_df)`` rows and ``r ncol(snp_df)`` columns. The variables in the dataframe are ``r names(snp_df)``.

Dataframe 3 `unemploy_df` has ``r nrow(unemploy_df)`` rows and ``r ncol(unemploy_df)`` columns. The variables in the dataframe are ``r names(unemploy_df)``.

The resulting dataset `merge3_df` is merged by the three above dataset according to the same year and month. It has ``r nrow(merge3_df)`` rows and ``r ncol(merge3_df)`` columns. The variables in the dataframe are ``r names(merge3_df)``. The range of years of the merged table is 
``r range(pull(merge3_df,year))[2]-range(pull(merge3_df,year))[1]`` years,from ``r range(pull(merge3_df,year))[1]`` to ``r range(pull(merge3_df,year))[2]``.


### Problem 3

















 
 


